---
title: "Day 6 Homework"
output: html_document
---
```{r}
library(reshape2)

```

1. Think carefully about your data.  Think about something you would like to do multiple times, to subsets of your data. 
-Just wanted to first take a look at the data again first to see what I'm working with...
```{r}
setwd("~/Dropbox/Winter 2015/R class")
d<- read.csv("AlisonS data.csv") #read in data
#head(d) #peak at data- need to cut out extra columns at end for now
d2<-d[ ,c(1:3,5, 10:16)]# just look at important cols
head(d2)
```

2. Write out (in plain English, in the notebook) what you are trying to do
For only the accurate trials (error = 0) I'd like to take the mean and standard deviation of each subject for future trimming purposes

3. Write some "pseudo-code" comments in an R chunk, as an outline
```{r}
# accdata<- mydata[mydata$error ==0] (or if mydata$ errror == 0...)
# for (sub1: numsubjects){
# get mean(RT), get SD(RT) and add to vectors meanRT, SD
```

4. Write some (real) code that will do what you want ONCE, on one subset of the data.
```{r}
d2$subject<-as.factor(d2$subject) #this was listed as an integer which is wrong
class(d2$subject) # this looks better- now I can subset by it
#if (d2$error ==0){
mysub<-d2[d2$subject == "101", ] #just sub101
my.meanRT<-mean(mysub$RT)
my.meanRT
```


5. Copy that code chunk, and convert it into a loop, which will do what you want over all the groups, concatenating the results.
```{r}
d2$subject<-as.factor(d2$subject)
my.subs<-c(101:139)
myRT.means<-vector(length = length(my.subs))
for(subnum in 1: length(my.subs)) {
mysub<-my.subs[subnum]#call nth sub
sub.RTs<-d2[d2$subject == mysub, "RT"]
myRT.means[subnum]<-mean(sub.RTs, na.rm = TRUE)
}
myRT.means #lists means and also missing subjects as Nan's
```

6. Copy that loop again, and at a condition (`if`) statement to add some further control about ignoring some groups or some other conditional operation.
```{r}
d2$subject<-as.factor(d2$subject)
my.subs<-c(101:139)
myRT.means<- rep("NA", length(my.subs))
for(subnum in 1: length(my.subs)) {
mysub<-my.subs[subnum]#call nth sub
sub.RTs<-d2[d2$subject == mysub, "RT"]
sub.errors<-d2[d2$subject == mysub, "error"]
#if (sub.errors[subnum] !=1) {   # GAH! I REALLY want to put this control in the loop (if error isn't 1, then continue) but I can't figure it out because the loop has to be finished to change the "subnum" to anything...I give up for now.
myRT.means[subnum]<-mean(sub.RTs, na.rm = TRUE)
}
}
myRT.means #lists means and also missing subjects as Nan's
```
Couldn't quite finish this last one the way I wanted to.
