---
title: "Day 7 Homework"
output: html_document
---

```{r}
library(reshape2)
library(ggplot2)
library(lme4)
library(stats)
getwd()
setwd("~/Dropbox/Winter 2015/R class")
d<- read.csv("AlisonS data.csv") #read in data
#head(d) #peak at data- need to cut out extra columns at end for now
d1<-d[ ,1:16]# just take the first 16 cols
head(d1)

d1$subject<-as.character(d1$subject) #code sub as character
d1$trial<-as.character(d1$trial) #code trial as character,
d1$trial<-as.factor(d1$trial) # then factor
d1$code<-as.character(d1$code) #code trial as character,
d1$trial<-as.factor(d1$code) # then factor
mycols <-c(1,2,5, 10:12) #narrow it down even more 
head(d1)
```

create log RT column: note- might want to log init.time as well?
```{r}
d1$logRT<-log(d1$RT) #log transform RT
summary(d1)
```
split up code into its components:

```{r}
dcode<-colsplit(d$code, "_", c("WordCond", "StroopCond", "half", "third")) # 
d2<-cbind(dcode, d1) #combine the new 4 cols with the old data- maybe delete old code col later...
head(d2) #
d2$half<-as.factor(d2$half)
d2$third<-as.factor(d2$third)
```
For looking at RT data- only look at subset of accurate data
```{r}
acc.d<-d2[d2$error ==0, ]
summary(acc.d$error) #check- yup- worked
```

Checking some cell means just for data exploration, tables, etc.
```{r}
#these use acc.d- only looking at accurate trials here
mean.RT<-dcast(acc.d, WordCond~StroopCond, value.var = "RT", fun.aggregate = ,mean, na.rm=TRUE)
mean.logRT<-dcast(acc.d, WordCond~StroopCond, value.var = "logRT", fun.aggregate = ,mean, na.rm=TRUE) #
mean.init.time<-dcast(acc.d, WordCond~StroopCond, value.var = "init.time", fun.aggregate = ,mean, na.rm=TRUE)
#looks like init.time might have something happening?
#error rate uses all data:
mean.error.rate<-dcast(d2, WordCond~StroopCond, value.var = "error", fun.aggregate = ,mean, na.rm=TRUE) # 

```
combine cols AUC1&2 and MD1&2 into 1 col each:
-*NOTE: THIS ISN"T WORKING-I've been mucking around with it a lot- says things about the replacement not being the right #..*l
```{r}
d2$AUC<- d2$AUC_2
length(d2$AUC)
length(d2$AUC_2)
d2$AUC[is.na(d2$AUC_2)] <-d2$AUC_2[is.na(d2$AUC_1)]
summary(d2$AUC) #new AUC column!
#same thing for MD
d2$MD<- d2$MD_2
head(d2)
d2$MD_1[is.na(d2$MD)] <-d2$MD_2[is.na(d2$MD_1)]
summary(d2$MD)
head(d2)
```
Split up "stimfile" column by \ and extract the .wav (and maybe also the final .jpg) to get the "item" (and maybe also to check if there is any influence of final Stroop color)
```{r}

```

Add in stroop error data and merge
```{r}

```

Add in LEAPQ data by subject and merge
```{r}

```

First contrast code factors:
```{r}
contrasts##finish here
```

Then run a preliminary analysis on logRT and init.time by word*stroop cond: (eventually will want to add a random slope by item as well)

```{r}
logRT.model<-lmer(logRT~WordCond*StroopCond+(WordCond+StroopCond|subject), data = d2)
summary(logRT.model) # no interactions
init.model<-lmer(log(init.time)~WordCond*StroopCond+(1|subject), data = d2) #log initiation time
summary(init.model) # no interactions - no effects of stroop...
```
analysis on errors per condition:
```{r}
error.model <-glmer(error~WordCond*StroopCond+(1|subject), data = d2, family = "binomial")
summary(error.model) # main effect of wordcond but not stroop or interaction
```
now same analysis but just look at 2nd/3rd third because the first part was kind of guessing.
```{r}
#note: for analyses the full models would not converge so I used reduced models...maybe try different optimizers?
end.d <-d2[d2$third != "t1",]
end.d<-droplevels(end.d)
summary(end.d)
logRT.model<-lmer(logRT~WordCond*StroopCond+(WordCond+StroopCond|subject), data = end.d)
summary(logRT.model) # now its very close to a main effect of stroop
init.model<-lmer(log(init.time)~WordCond*StroopCond+(WordCond+StroopCond|subject), data = end.d) #log initiation time
summary(init.model)# ooh here is an interesting ineraction: wordcondff*stroopcondinc t = -3.11 LOOK AT MEANS HERE
error.model<-glmer(error~WordCond*StroopCond+(1|subject), data = end.d, family = "binomial")
summary(error.model)
```

look at a plot of RT and init time time(thirds) by condition

```{r}
RT.p<-ggplot(d2, aes(StroopCond,logRT)) +geom_boxplot()
RT.p+ facet_wrap(~WordCond)

init.p<-ggplot(d2, aes(StroopCond,log(init.time))) +geom_boxplot()
               
init.p+facet_wrap(~WordCond)
```

