---
title: "Day 9 Homework"
output: html_document
---

1. Write a function to do a bunch of clean-up and formatting to your data.
```{r}
setwd("~/Dropbox/Winter 2015/R class")
d<- read.csv("AlisonS data.csv") #read in data
head(d)
colnames(d)

#remove errors FUNCTION!
LIMCA.clean<-function(dataframe, acc.col){
library(dplyr)
library(reshape2)
dataframe<-filter(dataframe, acc.col == 0)
droplevels(dataframe)
print(summary(dataframe))
dataframe$acc.col} ## This function was trying to do a lot more but I basically had to narrow it down to this to get it to work and didn't have time to build it back up.

```

2. Make sure this function works on your data.
```{r}
LIMCA.clean(dataframe= d, acc.col = "error") # I dont even think this is working anymore (thought it was but doesn't seem to be)
 summary(d$error)
```

3. Get some subset of your data, or find a similar-but-not-identical data set.  Change up the column order.  Re-run your function on this new data.  Does it work?
```{r}
newdata<-d[3:12,12:3]
head(newdata) # flip data around
LIMCA.clean(dataframe = newdata, acc.col = "error")
```

4. Fix your function to be able to work on the new data.
```{r}

  
```



5. Write a function to do some kind of summary stat, and use the `dplyr` functions to run this function over groups of your data.
```{r}
accrate<-function(dataframe, accCol){
    library(dplyr)
dataframe$subject<-as.factor(dataframe$subject)
dataframe%>%
  group_by(subject)%>%
  summarize (errorrate = mean(accCol, na.rm = TRUE))
}

acc.rate(d) #Error in library(., dplyr) : there is no package called ‘.’ ??
```


  
